---
title: "bioe215_debugging"
format: html
editor: visual
---

## Debugging with browser()

```{r}
library(palmerpenguins)
library(tidyverse)

traits_r2 <- function(trait1, trait2) {
  summary(lm(trait1 ~ trait2))$r.squared
}

penguins %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm))

```

**Q1** How would you describe the error? I\'m not asking you describe the *cause* of the error yet. Describe how the output is different than what you would expect.

There is no data output. It doesn't have r2 values

Q2: Where would you add browser() to explore the cause of the error?

I added it to the function before summary

Q3: Does the body of traits_r2 use list or atomic vector indexing? Does it use indexing by position, logic, or name?

Q4: What's the cause of the error? How would you fix it?

\$rsquared should be \$r.squared

```{r}
# Pipeline 1
penguins %>% 
  group_by(species) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm))

# Pipeline 2
penguins %>% 
  group_by(species, island) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm),
            .groups = "drop")

# Pipeline 3
penguins %>% 
  group_by(species, sex, island) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm),
            .groups = "drop")
```

Q5: How many times does Pipeline 1 call traits_r2()? How about Pipeline 2?

It calls it once for each species in pipe 1, so 3 times. In pipe 2, it runs 5 times for 3 species and 3 islands, but it's not 6 because not all species are on every island?

Q6: Create Pipeline 3 that additionally groups by sex. How many times does Pipeline 3 call traits_r2()?

It ran 13 times, including NAs in the sex
